---
title: "HR Attrition Neural Network"
output: html_document
date: "2022-12-19"
---

```{r}
library(keras)
library(tidyverse)
library(caret)
HRData <- read_csv("HR Employee Attrition.csv")
HRData <- HRData %>% select(c(-EmployeeCount, -StandardHours, -Over18, -EmployeeNumber))
HRData$Attrition<-ifelse(HRData$Attrition=="Yes",1,0)
# Update each column to be a factor – if it's a string 
#HRData <- mutate_if(HRData,is.character, factor)
```

Create test and train data sets
```{r}
trainList <- createDataPartition(y=HRData$Attrition, p=.70, list=FALSE)
trainData <- HRData[trainList, c("Attrition", "DistanceFromHome", "DailyRate", "Age", "Education")]
testData <- HRData[-trainList, c("Attrition", "DistanceFromHome", "DailyRate", "Age", "Education")]
```

Use a binary model
- units is the number of neurons
- ReLu is linear (identity) for positive values, and zero for negative
- "sigmoid" is typically used in the final layer for binary analysis
- It maps the output to a [0,1] range
- input_shape is the number of X inputs (we have 3)
```{r}
model = keras_model_sequential() %>%
  layer_dense(units=64, activation = "relu", input_shape=4) %>%
  layer_dense(units=1, activation="sigmoid")
```

Compile the model
- loss == how tyo measure error
- Use binary_crossentropy in binary problems where one of the two choices is correct
- optimizer == how to optimize during iterations
- optimizer_rmsprop is a good default
- metrics == list of metrics to be evaluated
- Often use metrics = "accuracy"
```{r}
compile(model,
        loss = "binary_crossentropy",
        optimizer=optimizer_rmsprop(),
        metrics="accuracy")
```

Run the binary model
- epochs: the number of times that the learning algorithm will work through the entire training dataset
- batch_size: the number of smaples to work through before updating the internal model parameters; 64, 128, 256 are typical default numbers
- validation_split: the percentage to validate the model - avoid overfitting
- the rest is to train the model
```{r}
x_data <- as.matrix(select(trainData, -Attrition))
history = fit(model,
              x_data,
              trainData$Attrition,
              epochs=20,
              batch_size=128,
              validation_split=0.2)
```

Use the model
- Do prediction
```{r}
x_data_test <- as.matrix(testData[,1:4])
y_data_pred=predict(model, x_data_test)
```

See how good the prediction was
```{r}
confusionMatrix(as.factor(y_data_pred), as.factor(testData$Attrition))
```