---
title: "HR Employee Analysis Data Exploration"
name: Sean Deery
output: pdf_document
date: "2022-11-07"
---

Read the CSV file into a dataframe
```{r}
library(tidyverse)
HRData <- read_csv("HR Employee Attrition.csv")
glimpse(HRData)
```

```{r}
range(HRData$EmployeeCount)
range(HRData$StandardHours)
table(HRData$Over18)
```

```{r}
HRData <- HRData %>% select(c(-EmployeeCount, -StandardHours, -Over18, -EmployeeNumber))
```


Attrition will be our dependent variable. 

We can get rid of EmployeeCount and EmployeeNumber because they won't contribute anything. We can also remove Over18 and StandardHours because they are the same for each employee.

The rest of the variables can be used as independent variables, and can be categorized as such:

- Personal Details: Age, Education, EducationField, Gender, MaritalStatus, NumCompaniesWorked, RelationshipSatisfaction

- Work Life: BusinessTravel, Department, DistanceFromHome, EnvironmentalSatisfaction, JobInvolvement, JobLevel, JobRole, JobSatisfaction, OverTime, PerformanceRating, TotalWorkingYears, TrainingTimesLastYear, WorkLifeBalance, YearsAtCompany, YearsInCurrentRole, YearsSinceLAstPromotion, YearsWithCurrentManager

- Income: DailyRate, HourlyRate, MonthlyIncome, MonthlyRate, PercentSalaryHike, StockOptionLevel


Check if there are any empty values.
```{r}
any(is.na(HRData))
```

Check out Attrition. 
```{r}
table(HRData$Attrition)
nrow(HRData[HRData$Attrition=="No",])/nrow(HRData)
nrow(HRData[HRData$Attrition=="Yes",])/nrow(HRData)
```
```{r}
ggplot(HRData) +
  aes(x=Attrition) +
  geom_bar()
```


Function to explore numerical columns
```{r}
exploreNumVariable <- function(num_variable) {
  # Get a summary of the data
  variable.summary <- summary(HRData[,num_variable])
  # Create the Group By
  variable.groupby <- HRData %>%
    group_by(Attrition) %>%
    summarize(var_mean=mean(.data[[num_variable]]))
  # Create the Histogram
  variable.hist <- ggplot(HRData) + 
    aes(x=.data[[num_variable]], fill=Attrition) +
    geom_histogram()
  # Create the Box Plot
  variable.box <- ggplot(HRData) + 
    aes(x=Attrition, y=.data[[num_variable]]) +
    geom_boxplot()
  return(list(variable.summary, variable.groupby, variable.hist, variable.box))
}
```

Function to explore categorical columns
```{r}
exploreCatVariable <- function(cat_variable) {
  # Create a table showing frequency distribution 
  variable.table <- table(HRData[, cat_variable])
  # Create a bar chart showing frequency distribution by Attrition
  variable.bar <- ggplot(HRData) +
    aes(x=.data[[cat_variable]], fill=Attrition) +
    geom_bar(position="dodge") +
    theme(axis.text.x=element_text(angle=90, hjust=1))
  return(list(variable.table, variable.bar))
}

```




Exploring Personal Details: 
- Age
- Education
- EducationField
- Gender
- MaritalStatus
- RelationshipSatisfaction

Age
- The average age of an employee who left is lower than the employees who don't.


```{r}
exploreNumVariable("Age")
```


Education
1 'Below College'
2 'College'
3 'Bachelor'
4 'Master'
5 'Doctor'
- It doesn;t seem like there is much of a difference when it comes to education levels.

```{r}
exploreNumVariable("Education")
```


EducationField

```{r}
exploreCatVariable("EducationField")
```


Gender

```{r}
exploreCatVariable("Gender")
```


MaritalStatus
- It looks like attrition is higher for employees who are single

```{r}
exploreCatVariable("MaritalStatus")
```

```{r}
single.attrition <- nrow(HRData[HRData$MaritalStatus=="Single" & HRData$Attrition=="Yes",])
single.total <- nrow(HRData[HRData$MaritalStatus=="Single",])
single.attrition/single.total

married.attrition <- nrow(HRData[HRData$MaritalStatus=="Married" & HRData$Attrition=="Yes",])
married.total <- nrow(HRData[HRData$MaritalStatus=="Married",])
married.attrition/married.total

divorced.attrition <- nrow(HRData[HRData$MaritalStatus=="Divorced" & HRData$Attrition=="Yes",])
divorced.total <- nrow(HRData[HRData$MaritalStatus=="Divorced",])
divorced.attrition/divorced.total
```



RelationshipSatisfaction
1 'Low'
2 'Medium'
3 'High'
4 'Very High'
- There doesn;t seem to be a difference in Attrition when it comes to relationship satisfaction

```{r}
exploreNumVariable("RelationshipSatisfaction")
```





Exploring Work Life: 
- BusinessTravel
- Department
- DistanceFromHome
- EnvironmentalSatisfaction
- JobInvolvement
- JobLevel
- JobRole
- JobSatisfaction
- NumCompaniesWorked
- OverTime
- PerformanceRating
- StandardHours
- TotalWorkingYears
- TrainingTimesLastYear
- WorkLifeBalance
- YearsAtCompany
- YearsInCurrentRole
- YearsSinceLAstPromotion
- YearsWithCurrentManager


BusinessTravel

```{r}
exploreCatVariable("BusinessTravel")
```


Department

```{r}
exploreCatVariable("Department")
```

```{r}
sales.attrition <- nrow(HRData[HRData$Department=="Sales" & HRData$Attrition=="Yes",])
sales.total <- nrow(HRData[HRData$Department=="Sales",])
sales.attrition/sales.total

rd.attrition <- nrow(HRData[HRData$Department=="Research & Development" & HRData$Attrition=="Yes",])
rd.total <- nrow(HRData[HRData$Department=="Research & Development",])
rd.attrition/rd.total

hr.attrition <- nrow(HRData[HRData$Department=="Human Resources" & HRData$Attrition=="Yes",])
hr.total <- nrow(HRData[HRData$Department=="Human Resources",])
hr.attrition/hr.total
```



DistanceFromHome

```{r}
exploreNumVariable("DistanceFromHome")
```


EnvironmentSatisfaction
1 'Low'
2 'Medium'
3 'High'
4 'Very High'

```{r}
exploreNumVariable("EnvironmentSatisfaction")
```


JobInvolvement
1 'Low'
2 'Medium'
3 'High'
4 'Very High'

```{r}
exploreNumVariable("JobInvolvement")
```


JobLevel
```{r}
exploreNumVariable("JobLevel")
```


JobRole
```{r}
exploreCatVariable("JobRole")
```


JobSatisfaction
1 'Low'
2 'Medium'
3 'High'
4 'Very High'
```{r}
exploreNumVariable("JobSatisfaction")
```


NumCompaniesWorked
```{r}
exploreNumVariable("NumCompaniesWorked")
```

OverTime
```{r}
exploreCatVariable("OverTime")
```


PerformanceRating
1 'Low'
2 'Good'
3 'Excellent'
4 'Outstanding'
```{r}
exploreNumVariable("PerformanceRating")
```


TotalWorkingYears
```{r}
exploreNumVariable("TotalWorkingYears")
```


TrainingTimesLastYear
```{r}
exploreNumVariable("TrainingTimesLastYear")
```


WorkLifeBalance
1 'Bad'
2 'Good'
3 'Better'
4 'Best'
```{r}
exploreNumVariable("WorkLifeBalance")
```


YearsAtCompany
```{r}
exploreNumVariable("YearsAtCompany")
```


YearsInCurrentRole
```{r}
exploreNumVariable("YearsInCurrentRole")
```



YearsSinceLastPromotion
```{r}
exploreNumVariable("YearsSinceLastPromotion")
```


YearsWithCurrManager
```{r}
exploreNumVariable("YearsWithCurrManager")
```


Exploring Income
- DailyRate
- HourlyRate
- MonthlyIncome
- MonthlyRate
- PercentSalaryHike
- StockOptionLevel


DailyRate

```{r}
exploreNumVariable("DailyRate")
```


HourlyRate

```{r}
exploreNumVariable("HourlyRate")
```


MonthlyIncome

```{r}
exploreNumVariable("MonthlyIncome")
```


MonthlyRate
```{r}
exploreNumVariable("MonthlyRate")
```


PercentSalaryHike
```{r}
exploreNumVariable("PercentSalaryHike")
```



StockOptionLevel
```{r}
exploreNumVariable("StockOptionLevel")
```

